On The Statistical Properties of Cospectra

### 2017-11-01 paper/referee_response.md

Response to the referee

We are grateful to the referee for their helpful and constructive comments. We believe that some of the issues raised by the referee may be result of a misunderstanding, and have realized that perhaps our text was not as clear as it should have been. We have expanded explanations in several cases, and have added a new section with caveats, as well as a data analysis on real NuSTAR data for illustrative purposes (all changes are marked in bold). Below, we lay out in detail our changes to the manuscript, and respond directly to the points the referee has raised. We hope that this helps solve the misunderstanding and satisfactorily addresses the referee's concern.

> Re. (i) Even for large Signal-to-noise (S/N) the co-spectrum is entirely insensitive to some oscillations with wrong phase. To my knowledge, no else known method suffers
> from such a weakness. Though remedies exist, the distributions discussed in the paper do not apply to them.
and

> 2. Re. (ii) and (iii) Fig. 3 reveals that the simulations were set up so, that the inserted oscillation remains in phase with the reference signal component, resulting with a large positive peak at the corresponding frequency. In the real word, the phases would be randomly distributed and both positive, negative and near null values would be observed.

The primary use case for the distributions defined in the manuscript are cases where observations were taken simulatenously of the same source with independent, identical detectors. This is now standard practice for multiple X-ray missions (including the now defunct RXTE, currently operating NuSTAR and AstroSAT, and future missions like IXPE). The practice of using the cospectrum of time series observed from two independent detectors, rather than the periodogram of the combined time series, was pioneered in NuSTAR as a means to mitigate the effects of dead time (and has since been used is >20 papers). However, the *underlying data stream* from the source is exactly the same. By definition, any periodic signal, be it in an X-ray pulsar or an accreting X-ray binary, must be strictly phase-aligned if it was observed simultaneously with detectors on the same spacecraft. We acknowledge that this might not be true for observations e.g. where the two time series were observed in different energy bands, since periods might be phase-shifted. We have clarified this point in the text at several occasions (in the Introduction, Section 2 as well as the new section "Caveats") and pointed out failure points.

Even in the latter cases, the cases we consider here always consider simultaneous observations of the same object, usually high-energy objects including pulsars, accreting neutron stars, or accreting black holes. While phase shifts are observed for different energy ranges, it is rather unlikely that the phases in the two observed time series are not covariant with each other, that is, while the *phase* itself might vary randomly if the underlying physical process is random, the phase *shift* will be set by specific (deterministic) conditions in the source (e.g. the geometry) and will therefore not be random.

> Re. (ii) The paper deals with a straw man set entirely in virtual reality.
> The quoted references indicate that the authors have access to real observations,
> yet make no attempt to analyze them with the proposed method.

As described above, the common practice of using the cospectrum derived from two identical detectors implies that the case discussed in the paper does, in fact, exist in reality. In particular, the primary reason this manuscript came into being is that we were both approached independently and asked how to define the statistical distributions for the cospectrum in the case of X-ray pulsation searches with NuSTAR. This, combined with the erroneous assumptions made about the statistical distributions used in the current astronomy literature on the topic (primarily the use of the Chi-square and normal distributions fairly indiscriminately), led us to believe that a manuscript laying out the basic statistical distributions to be used is warranted.

In response to the referee's comments, we have analyzed a set of pulsar observations with NuSTAR, and present the results in Section 4.2.

> Re. (iii) The authors make implicit assumptions which neither are obvious nor properly explained in the text. Thus the critical dependence of Co-spectrum on oscillation phase
> in not mentioned nor the fact that the only even-sampled observations are considered.
> This, and points (i) and (ii) are likely to confuse rather then enlighten a reader fresh in the field.

We have attempted to make our assumptions more explicit throughout the text, particularly in Section 2.

> 3. Re. (iii) Editing of Sect. 2.1 till Eq. (4) is so bad that it becomes source of major confusion for the reader. The text mentions a range of observation times without indication that it is implicitly assumed that observations are even-spaced. To demonstrate so suffice a counterexample of N(-) and N(+) of noise observations clustered around +/-T/2. Except for high frequencies, Fourier transform becomes [a exp(-i omega T/2) + b exp(+i omega T/2)] where Var(a)~1/N(-) differs from Var(b)~1/N(+), contrary to statements past Eq. (6).

We have attempted to edit Section 2.1 for clarity, and have explicitly stated that observations must be evenly sampled (and suggest alternative approaches if they are not).

> 4. A subtle flaw, common yet possibly serious of the derivations in the paper stems from assumption of known mean value 0 of x. This holds for simulations yet would be inappropriate for any real observations. Likely the central parts of the distributions would remain largely unaffected, yet their tails, important for the detection criteria, my be seriously affected possibly invalidating the said criteria. While precise accounting for this effect may be difficult, at least some limits to this spoiling effect should be provided.

We believe that there is a misunderstanding here. It is not the time series x that has a mean of zero, which indeed is unphysical, especially for photon counting data. Instead, the *Fourier amplitudes* have a mean of zero at a given frequency. This assumption will be true for any time series that is a realization of a wide-sense stationary stochastic process. For the simplest case of white noise, the Fourier amplitudes will be defined as A_j = \sum(x_k cos(omega_j*t_k)) and B_j = = \sum(x_k sin(omega_j*t_k)) where the sum is over k time intervals and omega_j is an angular frequency. If the x_k data points are distributed according to a normal distribution with mean \mu and variance \sigma^2 (where the mean could vary as a function of time, \mu = \mu_k), the distribution of the data points x_k with the sine or cosine will lead to a new normally distributed variable with a new mean of \mu_k*cos(omega_j*t_k) and a new variance of \mu_k*cos^2(omega_j*t_k). For frequencies between \nu_0 = 1/T up to the Nyquist frequency, this leads to values that are equally distributed around positive and negative means, and the distribution of the sum of all k values will not be exactly zero, but centred around zero. This will be true even when the process is not white noise, but a stationary "red noise" process, because the direction in either positive or negative direction of the time series around a mean is random, and thus the resulting Fourier amplitudes can tend to be either positive or negative.

This assumption underlies all of X-ray timing, and it is worth noting that in neither of us authors has encountered a case where this assumption has not been true, unless the time series was not a random process, not stationary, not evenly sampled or extremely sparse (with <100 photons in total or so). Thus, we conclude that for the vast majority of applications, this assumption will hold. We did add a statement in the relevant Section 2.1 after Equation 6 to make this more explicit.

> 5. while discussing signal from different detectors the authors do not mention an obvious option of adding power spectra to smooth out noise interference.

Averaging periodograms is a standard method to improve the signal-to-noise ratio in period searches. However, in this specific case of dead time, it is ineffective. Dead time occurs when after recording of a photon, it cannot detect another photon within a certain time span (it is, effectively, dead). This leads to a loss of overall photons, but also to a deviation from the typical Poisson process for photon arrival times to a process that involves a characteristic time scale. The resulting periodogram will show an oscillatory pattern over a large range of frequencies (depending on the characteristic time scale of the dead time), and changes both the mean and the variance of the distribution of powers as a function of frequency. Because the effect is always the same at a given frequency, averaging periodograms does not reduce the effect, but rather leads to narrow distributions with an unknown variance around an unknown mean. This is the primary motivation for the cospectrum as an alternative measure unaffected by dead time.

We have added a short description of the effects of dead time in the Introduction for clarification.

> 6. From the discussion in Sect 2.2.1 the reader could not guess that the
> Central Limit Theorem works here and the purpose of the section is mere check when it becomes good enough approximation.

We have added a half-sentence making that explicit.

> 7. While undoubtedly power spectrum is |F(nu)|^2, there exists a broad class of other estimates of PDF as functions of period/frequency. Hence e.g. wikipedia says: 'In signal processing, a periodogramm is an estimate of the spectral density of a signal ...Today, the periodogramm is a component of more sophisticated methods (see spectral estimation)'. Clearly a generic term is needed and periodogramm would fit nicely here.

Throughout the manuscript, we have tried to separate the different Fourier products quite clearly. As explained in Footnote 4 on page 1, we exclusively use the term "power spectrum" for the (stochastic) process generating a time series. Conversely, we exclusively use "periodogram" to denote the square of the absolute value of the Fourier amplitudes derived from a realization of the power spectrum, and say "cospectrum" or "cospectral powers" when we refer to the cospectral equivalent of the periodogram. In line with the signal processing literature, the cospectrum is explicitly not a subset of the periodogram, and we have kept those two concepts separate for clarity, since for understanding the paper it is crucial to make clear when we mean the squared Fourier amplitudes of a single time series, and when we mean the product of the Fourier amplitudes of one time series with the complex conjugate of the Fourier amplitude of a second time series.

### 2017-11-21 paper/cospectra-paper.tex
added two new figures with nustar results, associated notebook and text in paper draft

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author{D.~Huppenkothen\altaffilmark{1,2} and M.~Bachetti\altaffilmark{3}}

\altaffiltext{1}{Center for Data Science, New York University, 65 5h Avenue, 7th Floor, New York, NY 10003}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\altaffiltext{3}{INAF-Osservatorio Astronomico di Cagliari, via della Scienza 5, I-09047 Selargius (CA), Italy}

@@

@@

\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra. In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled. These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as X-ray timing of X-ray binaries, AGN and X-ray pulsar searches. Note that this assumes, however, that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers. Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}). Uneven sampling more generally introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid. For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}).}

@@

\section{Example: A QPO in \nustar}
\label{sec:nustarqpo}

@@

\subsection{Simulated Data}
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{qpo.png}
\caption{Comparison of the averaged periodograms from the two detectors (blue), the periodogram obtained by the sum of the light curves (yellow; both left panel) and the cospectrum (right panel), for $3000\,\mathrm{s}$ of synthetic \nustar data with an incident count rate of $200 \mathrm{counts}\mathrm{s}^{-1}$ and a strong QPO at $200\,\mathrm{Hz}$. 
The QPO has an rms of $\sim$9\% and a high Q factor of $\sim$40. The shape of the periodogram is strongly distorted from the expected flat powerspectrum centred on a constant value of $2$, with deviations of more that $0.5$ in the mean level of the powers common. In dashed lines, we also show the expected (trial-corrected) $99\%$ confidence level for the periodograms (in the same colour, left panel) as well as the cospectrum (red, right panel). The signal is detected in both the periodogram of the combined light curves as well as the cospectrum. However, because the white noise level is variable in the periodogram and the power spectrum dips below the expected level of $2$ around the frequency of the QPO, the signal is not detected as significantly as it should have been if no dead time was present. Assessing the significance is much more straightforward in the cospectrum, which retains a flat baseline.
}
\label{fig:qpo}
\end{center}
\end{figure*}

In order to show the difference of the detection limits with the cospectrum and the power spectrum, we show how a QPO at 200 Hz from a very bright source would appear in \nustar. We simulate a light curve of $T=3000\mathrm{s}$ duration with a time resolution of $\delta t = 0.5 \,\mathrm{ms}$ and an average count rate of $200 \,\mathrm{counts}\,\mathrm{s}^{-1}$. To this constant background we add a quasi-periodic oscillation with a period of $5\,\mathrm{ms}$, a fractional rms amplitude of $f_\mathrm{rms} = 0.15$ and phases randomized using a normal distribution with a mean of $0$ and a width of $\sigma_\mathrm{qpo} = 0.01$. We then simulated photon events using rejection sampling from this light curve using the software package \textit{stingray}\footnote{\url{https://github.com/StingraySoftware/stingray}}, running the function \texttt{stingray.Eventlist.simulate\_times()} twice in order to produce two light curves that are statistically independent, but have the same signal and properties, as we would expect from an instrument with two independent detectors observing the same object. Subsequently we simulated variable deadtime for both light curves with an average time scale of $2.5 \,\mathrm{ms}$ as commonly seen in \nustar\ data \citep{Bachetti+15} using \textit{HENDRICS}\footnote{\url{https://github.com/StingraySoftware/HENDRICS}} \citep{bachetti2015b}. We then produced the periodogram of the summed light curves, the averaged periodogram of the two individual light curves, and the cospectrum of the two light curves. Note that in all three cases we produced averaged periodograms and co-spectra by splitting the light curves into 600 segments of $5\,\mathrm{s}$ length each in order to suppress the variance in the powers and show the effects of dead time more clearly.

@@

@@

[ADD SOME STUFF ABOUT THE OBSERVATIONS.]

@@

The distributions laid out above allow for detecting periodic and narrow quasi-periodic signals in the presence of detector white noise, and especially important in the context of pulsar detection in X-rays, where faint sources may yield marginal detections even in the best of cases. At the same time, as instruments like \astrosat\ and \ixpe\ allow for observations with higher sensitivity, incorporating an accurate treatment of instrumental biases becomes increasingly important, and the cospectral statistics laid out here provide powerful tools to do so.
Notably absent from this discussion, however, is the much more common case where a source exhibits stochastic variability in the form of red noise or notably broadened quasi-periodic oscillations. In this case, the goal is either estimation of the precise properties of the underlying stochastic process, or detection of periods against a background varying stochastically. As has been shown above, the fact that the cross spectrum consists of two different time series complicates the statistical distributions considerably, and this is similarly true cospectra with variability. The exact treatment of this case is beyond the scope of this paper, and will be considered in depth in a forthcoming publication.

\paragraph{Acknowledgements}
The authors thank Thomas Laetsch for helpful suggestions regarding the mathematical derivations.
Daniela Huppenkothen is supported by the James Arthur Postdoctoral Fellowship and the Moore-Sloan Data Science Environment at New York University. 
MB is supported in part by the Italian Space Agency through agreement ASI-INAF n.2017-12-H.0 and ASI-INFN agreement n.2017-13-H.0.

>>>>

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author{D.~Huppenkothen\altaffilmark{1,2,3} and M.~Bachetti\altaffilmark{4}}

\altaffiltext{1}{Center for Data Science, New York University, 65 5h Avenue, 7th Floor, New York, NY 10003}
\altaffiltext{2}{Center for Cosmology and Particle Physics, Department of Physics, New York University, 4 Washington Place, New York, NY 10003, USA}
\altaffiltext{3}{DIRAC Institute, Department of Astronomy, University of Washington, 3910 15th Ave NE, Seattle, WA 98195}
\altaffiltext{4}{INAF-Osservatorio Astronomico di Cagliari, via della Scienza 5, I-09047 Selargius (CA), Italy}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{qpo.png}
\caption{Comparison of the averaged periodograms from the two detectors (blue), the periodogram obtained by the sum of the light curves (yellow; both left panel) and the cospectrum (right panel), for $3000\,\mathrm{s}$ of synthetic \nustar data with an incident count rate of $200 \mathrm{counts}\mathrm{s}^{-1}$ and a strong QPO at $200\,\mathrm{Hz}$. 
The QPO has an rms of $\sim$9\% and a high Q factor of $\sim$40. The shape of the periodogram is strongly distorted from the expected flat powerspectrum centred on a constant value of $2$, with deviations of more that $0.5$ in the mean level of the powers common. In dashed lines, we also show the expected (trial-corrected) $99\%$ confidence level for the periodograms (in the same colour, left panel) as well as the cospectrum (red, right panel). The signal is detected in both the periodogram of the combined light curves as well as the cospectrum. However, because the white noise level is variable in the periodogram and the power spectrum dips below the expected level of $2$ around the frequency of the QPO, the signal is not detected as significantly as it should have been if no dead time was present. Assessing the significance is much more straightforward in the cospectrum, which retains a flat baseline.
}
\label{fig:qpo}
\end{center}
\end{figure*}

@@

\textbf{All caveats applying more generally to Fourier analysis and to periodograms specifically also apply to cospectra. In particular, as for the $\chi^2$ distribution used for period searches in periodograms, the distributions derived here assume that the underlying process is wide-sense stationary and the light curves are simultaneously and evenly sampled. These conditions are typically met in most observations taken with X-ray timing instruments like \nustar\ and \rxte\ for problems such as timing of X-ray binaries, AGN and X-ray pulsar searches. Note that this assumes, however, that the time scales of interest are much shorter than outburst timescales in X-ray binaries or changes in the X-ray background in pulsar searchers. Non-stationarity (as is e.g.\ observed in flares) will impose a window function onto the cospectrum and shift the mean of the Fourier amplitudes, thus invalidating the assumptions above (for an illustrative example of this effect and its implications on period searches in periodograms of magnetar bursts, see \citealt{huppenkothen2013}). Uneven sampling more generally introduces covariances between adjacent frequencies and renders many of the statistical assumptions underpinning period searches using both in the periodogram and the cospectrum invalid. For unevenly sampled data, other methods that do not depend on a regular sampling pattern must be employed (for example the Lomb-Scargle periodogram \citep{lomb1976, scargle1982} or the Bayesian period search in \citet{gregory1992}).}

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{nustar_lc_psd.pdf}
\caption{\textbf{Upper panel: Light curve of Her X-1 observed with \nustar, observation [XX]. The observation totals $32.67\,\mathrm{ks}$, the light curve presented here is produced by combining event data from both FPMA and FPMB detectors. Gaps in the light curve are caused by the $96.8 \,\mathrm{m}$ orbit around Earth. Lower panel: averaged periodogram of $3260$ $10\,\mathrm{s}$-duration segments. The bright pulsar produces a very strong periodic signal at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$ ($P_\mathrm{rot} = 1.24\,\mathrm{s}$) and at the signal's harmonics. In order to show the effects of dead time on the underlying detector noise, we have zoomed into the noise region, such that the peaks produced by the pulsar extend beyond the upper edge of the plot. At the lowest frequencies, the periodogram deviates from the expected $\chi^2$ distribution centered on $\mu = 2$ because of longer-term, aperiodic variability in the source itself. At frequencies above a $\sim$5 Hz or so, dead time strongly modulates the noise powers into an oscillatory pattern, indicating that the use of standard statistical distributions commonly used to find periodic or quasi-periodic signals will lead to biased results.
}}
\label{fig:herx1_lc}
\end{center}
\end{figure*}

\section{Example: A (Q)PO in \nustar}
\label{sec:nustarqpo}

@@

\subsection{Simulated Data}
In order to show the difference of the detection limits with the cospectrum and the power spectrum, we show how a QPO at 200 Hz from a very bright source would appear in \nustar. We simulate a light curve of $T=3000\mathrm{s}$ duration with a time resolution of $\delta t = 0.5 \,\mathrm{ms}$ and an average count rate of $200 \,\mathrm{counts}\,\mathrm{s}^{-1}$. To this constant background we add a quasi-periodic oscillation with a period of $5\,\mathrm{ms}$, a fractional rms amplitude of $f_\mathrm{rms} = 0.15$ and phases randomized using a normal distribution with a mean of $0$ and a width of $\sigma_\mathrm{qpo} = 0.01$. We then simulated photon events using rejection sampling from this light curve using the software package \textit{stingray}\footnote{\url{https://github.com/StingraySoftware/stingray}}, running the function \texttt{stingray.Eventlist.simulate\_times()} twice in order to produce two light curves that are statistically independent, but have the same signal and properties, as we would expect from an instrument with two independent detectors observing the same object. Subsequently we simulated variable deadtime for both light curves with an average time scale of $2.5 \,\mathrm{ms}$ as commonly seen in \nustar\ data \citep{Bachetti+15} using \textit{HENDRICS}\footnote{\url{https://github.com/StingraySoftware/HENDRICS}} \citep{bachetti2015b}. We then produced the periodogram of the summed light curves, the averaged periodogram of the two individual light curves, and the cospectrum of the two light curves. Note that in all three cases we produced averaged periodograms and co-spectra by splitting the light curves into 600 segments of $5\,\mathrm{s}$ length each in order to suppress the variance in the powers and show the effects of dead time more clearly.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{nustar_cospectra.pdf}
\caption{\textbf{Upper panel: Cospectrum of a $10\,\mathrm{s}$ segment of the \nustar\ observations of Her X-1 (black). The cospectrum is created by observing the same source simultaneously, but statistically independently with the two identical FPMA and FPMB detectors. The rotational signal of the pulsar appears very clearly at $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$, but unlike the periodogram, the cospectrum is not affected by dead time. In red, we show the $p = 10^{-3}$ detection threshold, corrected for $6999$ trials (equivalent to the number of frequencies in the periodogram). Lower left:  histogram (grey) of the cospectral densities collected from 3260 individual cospectra between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$. In red, we also show the theoretically predicted Laplace distribution. Lower right: in black, the histogram of cospectral densities collected from $7$ averaged cospectra, each created from $15$ light curve segments of $200\,\mathrm{s}$ duration. Again, only cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ were used in this figure. In red, we show the theoretical probability density function defined in Equation \ref{eqn:averaged_pdf} with $n = 15$. For both averaged and single cospectra, the observed histogram matches the theoretically expected distributions very closely.
}}
\label{fig:herx1_cs}
\end{center}
\end{figure*}

@@

\begin{bf}
Hercules X-1 (Her X-1) is a well-studied persistent X-ray binary pulsars \citep{tananbaum1972} in an X-ray binary comprising the neutron star itself and a stellar companion HZ Herculis \citep{davidsen1972,forman1972,bahcall1972} with a mass of $\sim$$2.2 \,\mathrm{M}_\odot$ \citep{reynolds1997,leahy2014} whose type varies between late-type A and early-type B with orbital phase \citep{anderson1994,cheng1995} . The neutron star spins with a period of $P = 1.23\,\mathrm{s}$ [REF?] and the system overall exhibits an orbital period of $P_\mathrm{orb} = 1.7\,\mathrm{days}$ \citep{bahcall1972}, along with super-orbital variations on a $\sim$35-day timescale \citep{giacconi1973,scott1999,igna2011}.

Her X-1 was observed with the Nuclear Spectroscopic Telescope Array (\nustar) on XX for XX ks [REF!]. [ADD DETAILS ABOUT OBSERVATIONS AND DATA ANALYSIS HERE; SAY SOMETHING ABOUT NUSTAR DEAD TIME!]. Her X-1 was observed for a total of $32.67\,\mathrm{ks}$, counting only good time intervals longer than $10\,\mathrm{s}$. In Figure \ref{fig:herx1_lc} (upper panel), we present the light curve of the \nustar\ Her X-1 observations. The source varied substantially in brightness during the observations on fairly long time scales (10s to 1000s of seconds), indicating a significant source of red noise at low frequencies.

In the lower panel of Figure \ref{fig:herx1_lc}, we present an averaged periodogram created from a total of $3260$ light curve segments, each $10\,\mathrm{s}$ in duration. The periodogram shows very strong peaks at the $1.24\,\mathrm{s}$ rotational period of the pulsar, along with its first five harmonics. While at low frequencies, the deviation from the expected noise distribution (a $\chi^2$ distribution centered on $2$) can be explained with the longer-term variability within the source, at frequencies above $5 \,\mathrm{Hz}$ or so, the periodogram displays the typical oscillatory pattern associated with dead time (see also \citealt{Bachetti+15}), suggesting that the standard distributions usually applied to periodograms will not produce unbiased results for these observations.

In Figure \ref{fig:herx1_cs} (upper panel), we present a cospectrum of the Her X-1 \nustar\ observation. Because the periodic signal generated by the pulsar dominates the cospectrum, we plot the cospectrum of a short segment of $10\,\mathrm{s}$ duration starting at $\mathrm{MET} = 85740109.54169$ for illustrative purpose to highlight the noise properties. Unlike the periodogram, the cospectrum is not affected by dead time, and the statistical distributions defined in Section \ref{sec:whitenoise_cospectra} apply (see also Figure \ref{fig:herx1_cs}, lower left panel and below). We also show the trial-corrected detection threshold under the null hypothesis that the cospectrum consists solely of white noise. The highest cospectral density of $P = 55.16$ occurs at the frequency of the pulsar, $\nu_\mathrm{rot} = 0.806\,\mathrm{Hz}$; the probability of observing this power under the null hypothesis is effectively $p \sim 0$ within the limits of numerical accuracy.

In order to test our theoretical predictions for the distributions governing (averaged) cospectra, we first produced $3260$ individual cospectra out of each of the $10\,\mathrm{s}$ segments, and for each cospectrum extracted the cospectral densities in the range $50\,\mathrm{Hz}$ to $700\,\mathrm{Hz}$, where we do not expect significant contamination by the pulsar signal and its harmonics, or by the low-frequency red noise component.  In order to gain sufficient statistics, we plot a histogram of cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ for all $3260$ segments (a total of $21183500$ powers) in Figure \ref{fig:herx1_cs} (lower left panel), along with the theoretically expected Laplace distribution. We find a generally high agreement between the Laplace distribution and the distribution of observed cospectral densities: a Kolmogorov-Smirnov test yields $K= 1.77 \times 10^{-4}$ or a $p$-value of rejecting the null hypothesis that the observed distribution is derived from the theoretically expected Laplace distribution of $p = 0.52$.

In order to repeat the same process for averaged cospectra, we take the $7$ GTIs that are longer than $3000\,\mathrm{s}$  and extract a light curve of exactly $3000\,\mathrm{s}$ from each. From each of these light curves, we produce an averaged cospectrum by averaging $15$ consecutive light curve segments of $200\,\mathrm{s}$ duration. This yields $7$ averaged cospectra. As above, we extract cospectral densities between $50\,\mathrm{Hz}$ and $700\,\mathrm{Hz}$ and plot the histogram of cospectral densities extracted from all averaged cospectra (910000 powers) in Figure \ref{fig:herx1_cs} (lower right panel). As with the single cospectrum, the histogram for the averaged cospectral densities matches the theoretical prediction very closely.
\end{bf}

@@

\textbf{We show that the same is true for observations of the bright neutron star X-ray binary Her X-1, where the intrinsic brightness of the source leads to strong modulations in the periodogram due to dead time. For sources observed with \nustar\ and other instruments with a similar set-up of at least two redundant, identical detectors, one can take advantage of the independent, simultaneous observations in two detectors to form a cospectrum. While any (quasi-)periodic signal will appear weaker in the cospectrum than in the periodogram (even in the case, as presented here, where the signals in the two detectors are exactly in phase), it has the substantial advantage that it is unaffected by dead time and other similar detector effects. We show that the cospectrum of the observations observed with \nustar\ follow the expected theoretical distributions very closely.}

The distributions laid out above allow for detecting periodic and narrow quasi-periodic signals in the presence of detector white noise, and especially important in the context of pulsar detection in X-rays, where faint sources may yield marginal detections even in the best of cases. At the same time, as instruments like \astrosat\ and \ixpe, \textbf{with similar set-ups involving multiple identical detectors observing the same source simultaneously,} allow for observations with higher sensitivity, incorporating an accurate treatment of instrumental biases becomes increasingly important, and the cospectral statistics laid out here provide powerful tools to do so.
Notably absent from this discussion, however, is the much more common case where a source exhibits stochastic variability in the form of red noise or notably broadened quasi-periodic oscillations. In this case, the goal is either estimation of the precise properties of the underlying stochastic process, or detection of periods against a background varying stochastically. As has been shown above, the fact that the cross spectrum consists of two different time series complicates the statistical distributions considerably, and this is similarly true cospectra with variability. The exact treatment of this case is beyond the scope of this paper, and will be considered in depth in a forthcoming publication.

\paragraph{Acknowledgements}
The authors thank Thomas Laetsch for helpful suggestions regarding the mathematical derivations.
DH is supported by the James Arthur Postdoctoral Fellowship and the Moore-Sloan Data Science Environment at New York University. 
DH acknowledges support from the DIRAC Institute in the Department of Astronomy at the University of Washington. The DIRAC Institute is supported through generous gifts from the Charles and Lisa Simonyi Fund for Arts and Sciences, and the Washington Research Foundation.
MB is supported in part by the Italian Space Agency through agreement ASI-INAF n.2017-12-H.0 and ASI-INFN agreement n.2017-13-H.0.
